#' ---
#' title: Prevalence of heart failure and medical comorbidities in different electrocardiographic
#'   intraventricular conduction abnormalities compared to normal conduction
#' author: 
#'   - Daniel Loewenstein
#'   - Karin Johansson
#' date: '`r Sys.Date()`'
#' show_toc: yes
#' ---

## Summary

### Date

- Data extraction occured 2017-03-31 from [Deduce](https://www.ctsi.duke.edu/node/908)
  - [reference](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3063322/)

### Exclusions

- Exclude all patients with an endpointdate less than 24 hours after ecgdate.
- Exlucde all entries without EF data

## Setup

First we load all the needed packages.

```{r}
library(dplyr)
library(stringr)
library(purrr)
```

All the .xml ECG files have already been processed. You can read about the process [here](LP-xml_to_dataframe.html)

```{r}
#ecg <- DataPackageR::datapackager_object_read("ecg")
load(file.path(DataPackageR::project_data_path(), "ecg.rda"))
```

This is used for setting up a relative path during the package build.

```{r, eval=FALSE, echo=FALSE}
data_dir <- DataPackageR::project_data_path()
```

```{r}
# The above chunk is showed in vignette, but stored data otherwhere during build.

data_dir <- 
    file.path(path.expand("~"), "box", "jolo_projects", "promise", "inst", "extdata")
```

## Echo data

Time to read in all the echo data.

```{r, echo_data}
#Reads echo data
echo <- read.csv(file.path(data_dir, "all_echoes.csv"), 
                 stringsAsFactors = FALSE,
                na.strings = c(""," ","  ", "NA", "N/A"),
                strip.white = TRUE)
```

Since all the data are exported from the Deduce database, Duke University Medical Center, Durham, NC, USA we noticed that it's important to set the timezone (tz) explicitly, since some processing afterwards have been done outside NC and the default R settings is to use the local timezone.

```{r}
echo <- echo %>%
        rename(MRN = log_history, 
               EF = acq_efclosest,
               PAP.mm.hg = acq_rvsyspress,
               AS.severity = acq_as,
               MS.severity = acq_ms,
               AR.severity = acq_ai,
               MR.severity = acq_mr,
               echo_indication = log_sp1) %>%
        
        mutate(Echo.date = as.POSIXct(log_dos, 
                                      format = "%m/%d/%Y %H:%M:%S", 
                                      tz = "America/New_York"),
               PAP.mm.hg = as.numeric(PAP.mm.hg),
               EF.x = as.numeric(  #  had to remove characters to be able to convert to numericals
                   str_replace_all(EF, 
                                   c("<" = "", ">" = "", "%" = "", "=" = "")
                                   )
                   )
               ) %>%
        filter(!is.na(EF.x)) %>%  #  excludes all entries wihthout EF
        mutate_at(c("AS.severity",
                    "MS.severity",
                    "AR.severity",
                    "MR.severity"), as.factor) %>%
        select(MRN, 
               Echo.date, 
               EF, 
               EF.x, 
               AS.severity, 
               MS.severity, 
               AR.severity, 
               MR.severity, 
               PAP.mm.hg, 
               echo_indication)
               
        

```

## ECG and ECHO matching

We start by joining all patients by their Medical Record Number (MRN). We filter for all observations with less than or equal to 30 days, between the date of the echo and the date of the ecg. Among these we filter for the ones closest in time.

```{r, df}
# Joins echo and ecg by MRN variable
df <- inner_join(echo, ecg, by = "MRN")


df <- df %>%
        # group by MRN and echo date
        group_by(MRN) %>%

        # keeps observations with a difference between echo date and ecg date of 30 days or below
        filter(
            abs(
                as.numeric(
                    difftime(Echo.date, 
                             ECGDate, 
                             units = "days")
                    )
                )
            <= 30
            ) %>%
        
        # filters by echo exam and ecg exam closest in time (in seconds)
        filter(
            abs(
                as.numeric(
                    difftime(Echo.date,
                             ECGDate, 
                             units = "secs")
                    )
                ) 
            == min(
                abs(
                    as.numeric(
                        difftime(Echo.date, 
                                 ECGDate, 
                                 units = "secs")
                        )
                    )
                )
            ) %>%
        
        # removes eventual duplicates
        unique() %>%
        
        # ungroups to be able to do next operation
        ungroup() %>%
        
        # creates a new column holding number of NA in that row
        mutate(n = rowSums(is.na(.))) %>%
        
        # group by MRN
        group_by(MRN) %>%
        
        # within each MRN group selects the observation with the fewest numbers of NA
        filter(n == min(n)) %>%
        
        # for each MRN, select the first entry
        slice(1) %>%
        
        ungroup() %>%
        
        # drops the added column with number of na's
        select(-n)
```

Some patients when performing match in Duke DEDUCE database were flagged as having
gotten a new MRN. mrn.df contains a list of the old MRNs matched to the new ones.
However, 135 of the old MRNs exist in the supposedly new MRNs which we think are
due to that they had an ecg and echo exam both for the old MRN and the new MRN,
resulting in duplicates after the old MRN gets updated to the new MRN in our dataset.

Our solution to this is:

- identify which MRNs exist both as old and new, exclude these
- update the remaining old MRNs with the new numbers for further patient data mathing
which uses the updated MRNs.

```{r}
# Reads in updated MRN data
mrn.df <- read.csv(file.path(data_dir, "changedMrns.csv"), 
                   stringsAsFactors = FALSE,
                   strip.white = TRUE)

# sets columnnames
colnames(mrn.df) <- c("MRN", "new.MRN")

```

There were duplicates of the new MRN were different old MRNs had been assigned
the same new MRN, all of these are removed below.

```{r}
# selects all rows in mrn.df were new.MRN is not one of the new.MRNs that
# are duplicated.
mrn.df <- mrn.df[!mrn.df$new.MRN %in% 
                         mrn.df[duplicated(mrn.df$new.MRN),]$new.MRN,]

# Excludes all MRNs in df that also exist in mrn.df$new.MRN
df <- df[!df$MRN %in% mrn.df$new.MRN,]

```

Return all rows from df, and all columns from df and mrn.df. Rows in df with no match in mrn.df will have NA values in the new columns. If there are multiple matches between df and mrn.df, all combinations of the matches are returned.

```{r}
df <- left_join(df, mrn.df, by = "MRN")

# Replaces old MRNs with new ones, dropping the newly added new.MRN column.
df <- df %>%
        mutate(MRN = if_else(!is.na(new.MRN),
                             true = new.MRN, 
                             false = MRN)
               ) %>%
        select(-new.MRN)
```

## Demographics 

Reading in demographics datas regarding DOB, gender, race and date of death

```{r}
demo.df <- read.csv(file.path(data_dir, "DOB_GENDER_Race_DOD.csv"),
                    stringsAsFactors = FALSE,
                    na.strings = c(""," ","  ", "NA", "N/A"),
                    strip.white = TRUE)

```

Renaming column variables, setting date columns in POSIXct format, (POSIXlt uses a list format, while ct uses time in seconds from the origin, which is more memory effective), converting ordinal variables to factors.

```{r}
demo.df <- demo.df %>%
        transmute(MRN = Duke.MRN, 
               Gender = as.factor(Patient.Gender), 
               Race = as.factor(Patient.Race), 
               DOB = as.POSIXct(Patient.Date.of.Birth, 
                                format = "%m/%d/%Y %H:%M:%S", 
                                tz = "America/New_York"),
               Age.at.ECG = as.integer(0), # for future use
               Death.index = as.factor(Patient.Death.Indicator),
               DOD = as.POSIXct(Patient.Death.Date, 
                                format = "%m/%d/%Y %H:%M:%S", 
                                tz = "America/New_York"))
```


Let's clean up the Race factor levels a bit.

```{r}
levels(demo.df$Race) <- list("Mix"              = c("2 OR MORE RACES", 
                                                    "MULTIRACIAL"),
                             "American.native"  = c("ALASKAN NATIVE",
                                                     "AMERICAN INDIAN",
                                                     "AMERICAN INDIAN OR ALASKAN NATIVE"),
                             "Native.hawaiian.or.other.pacific.islander" = 
                               "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER",
                             "Caucasian"        = c("WHITE OR CAUCASIAN", 
                                                    "CAUCASIAN/WHITE"),
                             "African.american" = "BLACK OR AFRICAN AMERICAN",
                             "NA"               = c("UNAVAILABLE", 
                                                    "NOT REPORTED/DECLINED"),
                             "Asian"            = "ASIAN",
                             "Other"            = "OTHER")
```

## Matching demographics

This way of merging returns all rows from df where there are matching values in demo.df, and all columns from df and demo.df. If there are multiple matches between df and demo.df, all combination of the matches are returned.

```{r}
main.df <- inner_join(df, demo.df, by = "MRN") %>%
        
        # we need the patient age at the time of the ecg
        mutate(Age.at.ECG = round(
            as.integer(
                difftime(ECGDate,
                         DOB,
                         units = "days")
                )
            / 365.25)
            )
        
```


```{r}
#' Get dx file path
#'
#' @param file_name A \code{character}
#'
#' @return
#'
#' @examples
get_path <- function(file_name, sub_folder){
    path <- file.path(data_dir, paste0(sub_folder, "/", file_name))
}
```

```{r}
#' Import and format diagnosis data
#'
#' @param file_path \code{character}
#' @param abb_dx variable name, unquoted
#'
#' @return
#'
#' @examples
read_diagnosis_data <- function(file_name, abb_dx, type_data = c("diagnosis", "procedure")) {
    
    .sub_folder <- "Dx"
    .file_path <- get_path(file_name = file_name,
                            sub_folder = .sub_folder)
    .abb_dx <- enquo(abb_dx)
    .dx_code <- paste0(quo_name(.abb_dx), ".Code")
    .dx_date <- paste0(quo_name(.abb_dx), ".Date")
    
    .raw_data <- read.csv(
        .file_path,
        stringsAsFactors = FALSE,
        na.strings = c(""," ","  ", "NA", "N/A")
        )
    
    if (type_data[1] == "diagnosis") {
    .formatted_data <- .raw_data %>% 
        rename(MRN         = Duke.MRN,
               !!.dx_code := ICD.Diagnosis.Code,
               !!.dx_date := Diagnosis.Date) %>% 
        mutate(!!.dx_date := as.POSIXct(.data[[!!.dx_date]],
                                     format = "%m/%d/%Y %H:%M:%S",
                                     tz     = "America/New_York")
               )
    } else if (type_data[1] == "procedure") {
            .formatted_data <- .raw_data %>% 
        rename(MRN         = Duke.MRN,
               !!.dx_code := ICD.Procedure.Code,
               !!.dx_date := Procedure.Date) %>% 
        mutate(!!.dx_date := as.POSIXct(.data[[!!.dx_date]],
                                     format = "%m/%d/%Y %H:%M:%S",
                                     tz     = "America/New_York"),
               !!quo_name(.abb_dx) := as.factor(1)) %>%
                group_by(MRN) %>%
                filter(!!.dx_date == min(!!.dx_date)) %>%
                slice(1) %>%
                ungroup()
    }
}
```

### Procedure data

```{r}
crt.df <- read.csv(get_path("CRT Codes.csv", "Dx"), 
                   stringsAsFactors = FALSE,
                   na.strings = c(""," ","  ", "NA", "N/A")
                   )

crt.df <- crt.df %>%
        rename(MRN = Duke.MRN, 
               CRT.Date = ICD.CPT.Procedure.Date) %>%
        mutate(CRT.Date = as.POSIXct(CRT.Date, 
                                    format = "%m/%d/%Y %H:%M:%S", 
                                    tz = "America/New_York"),
               CRT = as.factor(1)
               ) %>%
        group_by(MRN) %>%
        filter(CRT.Date == min(CRT.Date)) %>%
        slice(1) %>%
        ungroup()
```

```{r}
lvad.df <- read_diagnosis_data(file_name = "LVAD Codes.csv", abb_dx = LVAD, type_data = "procedure")
htx.df <- read_diagnosis_data(file_name = "Hearttransplant Codes.csv", abb_dx = HTx, type_data = "procedure")
```


### Diagnosis data

```{r}
af.df <- read_diagnosis_data(file_name = "AF DX.csv", abb_dx = AF)
cad.df <- read_diagnosis_data(file_name = "CAD DX.csv", abb_dx = CAD)
chd.df <- read_diagnosis_data(file_name = "CHD DX.csv", abb_dx = CHD)
copd.df <- read_diagnosis_data(file_name = "COPD DX.csv", abb_dx = COPD)
dm.df <- read_diagnosis_data(file_name = "DM DX.csv", abb_dx = DM)
hf.df <- read_diagnosis_data(file_name = "HF DX.csv", abb_dx = HF)
htn.df <- read_diagnosis_data(file_name = "HTN DX.csv", abb_dx = HTN)
osa.df <- read_diagnosis_data(file_name = "OSA DX.csv", abb_dx = OSA)
```


Creating new dataframe containing just the MRNs and the corresponding ECGdates
so that this can be used in the dx.match function to match diagnosis date to ecg date.

```{r}
ecgdates.df <- main.df %>%
        select(MRN, ECGDate)
```

Function to select diagnosis meeting match criteria, the first date that is not more than >30 days after ECG date.

For some patients there will be multiple entries with the same diagnosis date, this is due to different diagnosis codes, which one does not matter so we are selecting the first one.

```{r}
#' Match MRN to diagnosis data
#'
#' @param x \code{data.frame} with diagnosis data.
#' @param diagnosis abbreviated diagnosis, unquoted.
#'
#' @return
#'
#' @examples
match_mrn_diagnosis <- function(x, diagnosis) { 
  
    .dx <- enquo(diagnosis)
    .dx_date <- paste0(quo_name(.dx), ".Date")
        
    .matched_data <- x %>% 
        inner_join(ecgdates.df, by = "MRN") %>%  #  join x with ecgdates dataframe
        group_by(MRN) %>% 
        filter(!!.dx_date == min(!!.dx_date)) %>%
        filter(as.numeric(
            difftime(ECGDate,
                     .data[[!!.dx_date]],
                     units = "days")
        )
        >= (-30)
        ) %>% 
        slice(1) %>% 
        ungroup() %>% 
        mutate(!!quo_name(.dx) := factor(1)) %>%  #  Diagnosis indicator
        select(-ECGDate)
}
```

```{r}
af.df <- match_mrn_diagnosis(af.df, AF)
cad.df <- match_mrn_diagnosis(cad.df, CAD)
chd.df <- match_mrn_diagnosis(chd.df, CHD)
copd.df <- match_mrn_diagnosis(copd.df, COPD)
dm.df <- match_mrn_diagnosis(dm.df, DM)
hf.df <- match_mrn_diagnosis(hf.df, HF)
htn.df <- match_mrn_diagnosis(htn.df, HTN)
osa.df <- match_mrn_diagnosis(osa.df, OSA)

```

Joining diagnosis data to main.df

```{r}

combine_list <- list(main.df, 
                     af.df, 
                     cad.df, 
                     chd.df, 
                     copd.df, 
                     dm.df, 
                     hf.df, 
                     htn.df, 
                     osa.df,
                     crt.df,
                     lvad.df,
                     htx.df)

main.df2 <- purrr::reduce(combine_list, left_join, by = "MRN")

```

```{r}
# Exclude all patients with an endpointdate less than 24 hours after ecgdate.

endPointAfterEcg <- function(ecgdate, endpointdate) {
  x <- (as.numeric(difftime(ecgdate, endpointdate, units = "days")) <= (-1))
}


main.df2 <- main.df2 %>%
  group_by(MRN) %>%
  filter((endPointAfterEcg(ECGDate, CRT.Date) | is.na(CRT.Date))) %>%
  filter((endPointAfterEcg(ECGDate, LVAD.Date) | is.na(LVAD.Date))) %>%
  filter((endPointAfterEcg(ECGDate, HTx.Date) | is.na(HTx.Date))) %>%
  ungroup()
```


## Medications

```{r}
read_rx_data <- function(file_name, abb_rx, patient = c("in", "out")){
    
    .rx <- enquo(abb_rx)
    .path <- get_path(file_name = file_name,
                      sub_folder = "Medications")
    
    .rx_data <- read.csv(.path,
                         stringsAsFactors = FALSE,
                         na.strings = c(""," ","  ", "NA", "N/A"))
    
    if (patient[1] == "in") {
        .rx_name <- paste0(quo_name(.rx), ".Rx.Name")
        .rx_start_date <- paste0(quo_name(.rx), ".Rx.Start")
        .rx_end_date   <- paste0(quo_name(.rx), ".Rx.End")
        
        .formatted_data <- .rx_data %>% 
            rename(MRN = Duke.MRN,
                   !!.rx_name := Medication.Name,
                   !!.rx_start_date := Start.Date,
                   !!.rx_end_date   := End.Date) %>% 
            mutate(!!.rx_start_date := as.POSIXct(.data[[!!.rx_start_date]],
                                                  format = "%m/%d/%Y %H:%M:%S", 
                                                  tz = "America/New_York"),
                   !!.rx_end_date   := as.POSIXct(.data[[!!.rx_end_date]],
                                                  format = "%m/%d/%Y %H:%M:%S", 
                                                  tz = "America/New_York")
            )
    } else if (patient[1] == "out") {
        .rx_name <- paste0(quo_name(.rx), ".Rx.OMR.Name")
        .rx_date <- paste0(quo_name(.rx), ".Rx.OMR.Date")
        
        
        .formatted_data <- .rx_data %>% 
            rename(MRN = Duke.MRN,
                   !!.rx_name := OMR.Medication.Name,
                   !!.rx_date := OMR.Report.Date) %>% 
            mutate(!!.rx_date := as.POSIXct(.data[[!!.rx_date]],
                                            format = "%m/%d/%Y %H:%M:%S", 
                                                  tz = "America/New_York")
            )
    }
}
```


```{r}
classIII.df <- read_rx_data("Class III Patient Medications.csv", ClassIII)
classIc.df <- read_rx_data("Class Ic Patient Medication data.csv", ClassIc)

bb_list <- list("BB Patient Med2013 to 2017.csv",
                "BB Patient Med2010 to 2013.csv",
                "BB Patient Med2006 to 2010.csv")
bb.df <- map_df(bb_list, read_rx_data, abb_rx = BB)

amiodarone.df <- read_rx_data("AmiodaronePatient Medication data.csv",Amiodarone)

acei_list <- list("ACEI Patient Medications 2013to2017.csv",
                  "ACEI Patient Medications 2010to2013.csv",
                  "ACEI Patient Medications 2006to2010.csv")

acei.df <- map_df(acei_list, read_rx_data, abb_rx = ACEI)
```


We only want medication entries where the start date was not more than 30 days after ecgdate,
or where the end date was more than 30 days before ecgdate.

For outpatients, denoted by OMR, where we only have a date indicating whether that patient had the medication at that point in time or not, we filter for the absolute difference in days between that timepoint and ecgdate to be less than or equal to 30 days.

```{r}
match_mrn_rx <- function(x, rx, patient = c("in", "out")){
    .rx <- enquo(rx)
    .matched_data <- inner_join(ecgdates.df, x, by = "MRN")
    .rx_name <- paste0(quo_name(.rx), ".Rx")
    
    if (patient[1] == "in"){
        .rx_start_date <- paste0(quo_name(.rx), ".Rx.Start")
        .rx_end_date <- paste0(quo_name(.rx), ".Rx.End")
        
        .formatted_data <- .matched_data %>% 
            mutate(
                start_diff = as.numeric(
                    difftime(ECGDate, .data[[!!.rx_start_date]], units = "days")
                ),
                end_diff = as.numeric(
                    difftime(ECGDate, .data[[!!.rx_end_date]], units = "days")
                )
            ) %>% 
            group_by(MRN) %>% 
            filter(start_diff >= (-30)) %>%
            filter(end_diff <= 30) %>% 
            slice(1) %>% 
            ungroup() %>% 
            mutate(!!.rx_name := factor(1)) %>% 
            select(-ECGDate, -start_diff, -end_diff)
    } else if (patient[1] == "out") {
        .rx_report_date <- paste0(quo_name(.rx), ".Rx.OMR.Date")
        
        .formatted_data <- .matched_data %>% 
            group_by(MRN) %>%
            mutate(time_diff = abs(
                as.numeric(
                    difftime(ECGDate,
                             .data[[!!.rx_report_date]],
                             units = "days")
                    )
                )
            ) %>% 
            filter(time_diff == min(time_diff, na.rm = TRUE)) %>% 
            filter(time_diff <= 30) %>% 
            slice(1) %>% 
            ungroup() %>% 
            mutate(!!.rx_name := factor(1)) %>%
            select(-time_diff, -ECGDate)
    }
}
```

```{r}
classIII.df <- match_mrn_rx(classIII.df, ClassIII)
classIc.df <- match_mrn_rx(classIc.df, ClassIc)
acei.df <- match_mrn_rx(acei.df, ACEI)
amiodarone.df <- match_mrn_rx(amiodarone.df, Amiodarone)
bb.df <- match_mrn_rx(bb.df, BB)
```

```{r}
classIII.OMR.df <- read_rx_data("Class III OMR data.csv", ClassIII, patient = "out")
classIc.OMR.df <- read_rx_data("Class Ic OMR data.csv", ClassIc, patient = "out")

bb_OMR_list <- list(
                "BB OMR2010 to 2013.csv",
                "BB OMR2006 to 2010.csv"
                )
bb.OMR.df <- map_df(bb_OMR_list, read_rx_data, abb_rx = BB, patient = "out")

amiodarone.OMR.df <- read_rx_data("AmiodaroneOMR data.csv", Amiodarone, patient = "out")

acei_OMR_list <- list(
                  "ACEI OMR2010 to 2013.csv",
                  "ACEI OMR2006 to 2010.csv"
                  )

acei.OMR.df <- map_df(acei_OMR_list, read_rx_data, abb_rx = ACEI, patient = "out")
```

```{r}
classIII.OMR.df <- match_mrn_rx(classIII.OMR.df, ClassIII, patient = "out")
classIc.OMR.df <- match_mrn_rx(classIc.OMR.df, ClassIc, patient = "out")
acei.OMR.df <- match_mrn_rx(acei.OMR.df, ACEI, patient = "out")
amiodarone.OMR.df <- match_mrn_rx(amiodarone.OMR.df, Amiodarone, patient = "out")
bb.OMR.df <- match_mrn_rx(bb.OMR.df, BB, patient = "out")
```
Joining diagnosis data to main.df

Earlier added column indicating whether specific type of medication was used or not except from joining hospital and OMR patient medcation data this code also updates that column with a 1 if presence of entry in OMR medication column

```{r}
combine_rx_list <- list(
    main.df2,
    classIII.df,
    classIc.df,
    acei.df,
    amiodarone.df,
    bb.df,
    classIII.OMR.df,
    classIc.OMR.df,
    acei.OMR.df,
    amiodarone.OMR.df,
    bb.OMR.df
)


main.df3 <- purrr::reduce(combine_rx_list, left_join, by = "MRN")

```

```{r}
#' Coalesce duplicated Rx columns
#'
#' The matching of MRN and Rx created .Rx.x and .Rx.y indicator columns, this merges those.
#'
#' @param abb_rx A \code{character} Rx name
#' @param x A \code{data.frame}
#'
#' @return
#'
#' @examples
coalesce_rx_columns <- function(abb_rx, x){
    .rx_x <- paste0(abb_rx, ".Rx.x")
    .rx_y <- paste0(abb_rx, ".Rx.y")
    .rx_name <- paste0(abb_rx, ".Rx")
    
    .formatted_data <- x %>% 
        mutate(!!.rx_name := coalesce(.data[[!!.rx_x]], 
                                      .data[[!!.rx_y]])
               ) %>% 
        select(-!!.rx_x, -!!.rx_y)
}
```


```{r}
coalesce_list <- list(
    "ClassIII",
    "ClassIc",
    "ACEI",
    "Amiodarone",
    "BB")

for(rx in coalesce_list){
    main.df3 <- coalesce_rx_columns(rx, main.df3)
}
```


## Lab data

```{r}
read_lab_data <- function(file_name, lab) {
    .lab <- enquo(lab)
    .lab_date <- paste0(quo_name(.lab), ".lab.Date")
    .lab_value <- paste0(quo_name(.lab), ".lab.Value")
    
    .path <- get_path(file_name = file_name,
                      sub_folder = "Lab")
    
    .lab_data <- read.csv(.path,
                          stringsAsFactors = FALSE,
                          na.strings = c(""," ","  ", "NA", "N/A", 
                                         "CANCELED", "*H", "*", "**", 
                                         "***", "*******")
                          )
    
    .formatted_data <- .lab_data %>% 
        rename(MRN = Duke.MRN,
               !!.lab_date := Specimen.Collection.Date,
               !!.lab_value := Text.Result) %>% 
        mutate(!!.lab_date := as.POSIXct(.data[[!!.lab_date]],
                                         tryFormats = c(
                                             "%m/%d/%Y %H:%M:%S",
                                             "%m/%d/%Y %H:%M"),
                                         tz = "America/New_York")
        ) %>% 
        # lab values contains strings like > < for greater, lower than, also , as thousand marker
        # remove for coercion to numeric.
        mutate(!!.lab_value := stringr::str_remove_all(.data[[!!.lab_value]], pattern = "<|>|=|,"))
    
    missing_text_result_index <- which(is.na(.formatted_data[[.lab_value]]) &
                                     !is.na(.formatted_data[["Numeric.Result"]]))
    
    replacement_values <- as.numeric(.formatted_data[missing_text_result_index, "Numeric.Result"])
    
    .formatted_data[missing_text_result_index, .lab_value] <- replacement_values
    
    if (quo_name(.lab) == "BNP") {
        
        bnp_levels <- list(
            "BNP - LABCORP" = "BNP",
            "BRAIN NATRIURETIC PEPTIDE" = "BNP",
            "BRAIN NATRIURETIC PEPTIDE     (BKR)" = "BNP",
            "BTYPE NATRIURETIC PEPTIDE DUAP" = "BNP",
            "PRO BRAIN NATRIURETIC PEPTIDE" = "ProBNP",
            "PRO BRAIN NATRIURETIC PEPTIDE     (BKR)" = "ProBNP",
            "PROBNP - LABCORP" = "ProBNP"
        )
        
        .formatted_data <- .formatted_data %>% 
            select(-Numeric.Result) %>% 
            filter(!is.na(!!.lab_value)) %>% 
            mutate(Test.Description = as.factor(Test.Description)) %>% 
            mutate(Test.Description = recode_factor(Test.Description,
                                                    !!!bnp_levels)
            )
    } else {
        .formatted_data <- .formatted_data %>% 
            select(-Test.Description, -Numeric.Result) %>% 
            filter(!is.na(!!.lab_value))
    }
}
```

```{r}
na_lab_list <- list("NA2013_2016.csv",
                    "NA2010_2013.csv",
                    "NA2006_2010.csv",
                    "NA2001_2006.csv")

na.df <- purrr::map_df(na_lab_list, read_lab_data, lab = Na)
```

```{r}
k_lab_list <- list("K2013_2016.csv",
                   "K2010_2013.csv",
                   "K2006_2010.csv",
                   "K2001_2006.csv")

k.df <- purrr::map_df(k_lab_list, read_lab_data, lab = K)
```

```{r}
gfr.df <- read_lab_data("GFR DX.csv", GFR)
```

```{r}
bnp_pro_bnp.df <- read_lab_data("BNP AND PRO BNP2.csv", BNP)
```

```{r}
bnp.df <- bnp_pro_bnp.df %>%
        filter(Test.Description == "BNP") %>%
        select(-Test.Description)

probnp.df <- bnp_pro_bnp.df %>%
        filter(Test.Description == "ProBNP") %>%
        rename(ProBNP.lab.Date = BNP.lab.Date,
               ProBNP.lab.Value = BNP.lab.Value) %>%
        select(-Test.Description)
```

- Filters for the labvalue closest in time to ecgdate
- Filters for the lab date that is within 30 days of ecgdate

```{r}
match_mrn_lab <- function(x, lab) {
    .lab <- enquo(lab)
    
    # Turn .lab_date and .lab_value into symbols for use with dplyr
    .lab_date <- rlang::sym(paste0(quo_name(.lab), ".lab.Date"))
    .lab_value <- rlang::sym(paste0(quo_name(.lab), ".lab.Value"))
    
    
    .matched_data <- inner_join(ecgdates.df, x, by = "MRN")
    
    
    .formatted_data <- .matched_data %>% 
        group_by(MRN) %>%
        mutate(time_diff = 
                   abs(
                       as.numeric(
                           difftime(
                               ECGDate,
                               !!.lab_date,
                               units = "days")
                       )
                   )
        ) %>% 
        filter(time_diff == min(time_diff, na.rm = TRUE)) %>%
        filter(time_diff <= 30) %>%
        group_by(MRN, ECGDate, !!.lab_date, time_diff) %>% 
        # Some patients have multiple lab values/day, return the mean
        summarize(!!.lab_value := mean(as.numeric(!!.lab_value), na.rm = TRUE)) %>% 
        slice(1) %>% 
        ungroup() %>% 
        select(-time_diff, -ECGDate)
}
```

```{r}
na.df <- match_mrn_lab(na.df, Na)
k.df <- match_mrn_lab(k.df, K)
bnp.df <- match_mrn_lab(bnp.df, BNP)
probnp.df <- match_mrn_lab(probnp.df, ProBNP)
gfr.df <- match_mrn_lab(gfr.df, GFR)
```

```{r}
combine_lab_list <- list(
    main.df3,
    na.df,
    k.df,
    bnp.df,
    probnp.df,
    gfr.df
)

main.df4 <- purrr::reduce(combine_lab_list, left_join, by = "MRN")
```

## Data processing

### Exclusion criteria

- Patients under 18 years of age
- QRSduration <60ms

```{r}

main.df4 <- main.df4 %>%
        
        filter(Age.at.ECG >= 18,
               
               QRSDur >= 60)
```

```{r}
main.df5 <- main.df4 %>%
        mutate(RBBB = if_else(RBBB == 1 & QRSDur >= 120,
                               true  = 1,
                               false = 0),
               
               RBBB_LAFB = if_else(RBBB == 1 & LAFB == 1 & (AXIS <= -30 & AXIS >= -89),
                                    true  = 1,
                                    false = 0),
               
               RBBB_LPFB = if_else(RBBB == 1 & LPFB == 1 & (AXIS >= 110 & AXIS <= 180),
                                    true  = 1,
                                    false = 0),
               
               RBBB = if_else(RBBB_LAFB == 1 | RBBB_LPFB == 1,
                               true  = 0,
                               false = RBBB),
               
               LBBB = if_else(LBBB == 1 & QRSDur >= 120,
                               true  = 1,
                               false = 0),
               
               IVCD = if_else(IVCD == 1 & QRSDur >= 110 &
                                       LBBB == 0 &
                                       RBBB == 0 &
                                       RBBB_LAFB == 0 &
                                       RBBB_LPFB == 0,
                               true  = 1,
                               false = 0),
               
               LAFB = if_else(LAFB == 1 & QRSDur <120 & IVCD == 0 & (AXIS <= -30 & AXIS >= -89),
                               true  = 1,
                               false = 0),
               
               LPFB = if_else(LPFB == 1 & QRSDur <120 & IVCD == 0 & (AXIS >= 110 & AXIS <= 180),
                               true  = 1,
                               false = 0),
               
               Normal_conduction = if_else(LBBB      == 0 & 
                                           LAFB      == 0 &
                                           LPFB      == 0 &
                                           RBBB      == 0 &
                                           RBBB_LAFB == 0 &
                                           RBBB_LPFB == 0 &
                                           IVCD      == 0 &
                                           PACE      == 0 &
                                           QRSDur < 120,
                                           
                                           true  = 1,
                                           false = 0))
               
               
```

```{r}
replace_na <- function(x) {
        x[is.na(x)] <- 0
        x
}
```

```{r}
fix_na_columns <- c("AF", 
                    "CAD", 
                    "CHD",
                    "COPD",
                    "CRT",
                    "DM",
                    "HF",
                    "HTN",
                    "HTx",
                    "LVAD",
                    "OSA",
                    "ClassIII.Rx",
                    "ClassIc.Rx",
                    "ACEI.Rx",
                    "Amiodarone.Rx",
                    "BB.Rx")
```


```{r}
main.df5 <- main.df5 %>%
        mutate_at(fix_na_columns,as.numeric) %>%
        mutate_at(fix_na_columns,replace_na)
```

```{r}
# patients without ecg classification or multiple entries are excluded
main.df6 <- main.df5 %>%
        rowwise() %>%
        mutate(n_entries = sum(Normal_conduction,
                               IVCD,
                               LBBB,
                               LAFB,
                               LPFB,
                               RBBB,
                               RBBB_LAFB,
                               RBBB_LPFB,
                               PACE)) %>%
        filter(n_entries == 1) %>%
        select(-n_entries) %>%
        ungroup()
```

```{r}
promiseData <- droplevels(main.df6)
```



