---
title: "Prevalence of heart failure and medical comorbidities in different electrocardiographic intraventricular conduction abnormalities compared to normal conduction"
author: "Daniel Loewenstein"
date: "`r Sys.Date()`"
slug: promise
show_toc: yes
output:
  knitr:::html_vignette: ~
  number_sections: yes
vignette: >
  %\VignetteIndexEntry{Reading xml ECG and processing diagnosis}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---


## Setup

First we load all the needed packages.

```{r}
library(dplyr)
library(readr)
library(lazyeval)
library(stringr)
library(xml2)
library(purrr)
library(data.table)
```

All the .xml ECG files have already been processed. You can read about the process [here](LP-xml_to_dataframe.html)

```{r}
ecg <- readRDS(file.path(DataPackageR::project_path(), "data-raw/promise_xml_ecg-GEN.rds"))
```

This is used for setting up a relative path during the package build.

```{r}
data_dir <- "/mnt/c/Users/Daniel/Cloudstation/Documents/Research/Projects/POHMECG/data/"
```

## Echo data

Time to read in all the echo data.

```{r, echo_data}
#Reads echo data
echo <- read.csv(file.path(data_dir, "all_echoes.csv"), 
                 stringsAsFactors = FALSE,
                na.strings=c(""," ","  ", "NA", "N/A"),
                strip.white = TRUE)
```

Since all the data are exported from the Deduce database, Duke University Medical Center, Durham, NC, USA we noticed that it's important to set the timezone (tz) explicitly, since some processing afterwards have been done outside NC and the default R settings is to use the local timezone.

```{r}
#Renames Echo.MRN to MRN
echo <- echo %>%
        rename(MRN = log_history, 
               EF = acq_efclosest,
               PAP.mm.hg = acq_rvsyspress,
               AS.severity = acq_as,
               MS.severity = acq_ms,
               AR.severity = acq_ai,
               MR.severity = acq_mr,
               echo_indication = log_sp1) %>%
        
        mutate(Echo.date = as.POSIXct(log_dos, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
               PAP.mm.hg = as.numeric(PAP.mm.hg),
               
               #had to remove characters to be able to convert to numericals
               EF.x = as.numeric(str_replace_all(EF, c("<" = "", ">" = "", "%"= "", "=" = "")))) %>%
        
        #excludes all entries wihthout EF
        filter(!is.na(EF.x)) %>%
        
        mutate_at(c("AS.severity",
                    "MS.severity",
                    "AR.severity",
                    "MR.severity"), as.factor) %>%
        
        select(MRN, Echo.date, EF, EF.x, AS.severity, MS.severity, AR.severity, MR.severity, PAP.mm.hg, echo_indication)
               
        

```

## ECG and ECHO matching

We start by joining all patients by their Medical Record Number (MRN). We filter for all observations with less than or equal to 30 days, between the date of the echo and the date of the ecg. Among these we filter for the ones closest in time.

```{r, df}
#Joins echo and ecg by MRN variable
df <- inner_join(echo, ecg, by = "MRN")


df <- df %>%
        #group by MRN and echo date
        group_by(MRN) %>%

        #keeps observations with a difference between echo date and ecg date of 30 days or below
        filter(abs(as.numeric(difftime(Echo.date,ECGDate, units = "days"))) <= 30) %>%
        
        #filters by echodate for the observations closest in time (in seconds)
        filter(abs(as.numeric(difftime(Echo.date,ECGDate, units = "secs"))) == min(abs(as.numeric(difftime(Echo.date, ECGDate, units = "secs"))))) %>%
        
        #removes eventual duplicates
        unique() %>%
        
        #ungroups to be able to do next operation
        ungroup() %>%
        
        #creates a new column holding number of NA in that row
        mutate(n = rowSums(is.na(.))) %>%
        
        #group by MRN
        group_by(MRN) %>%
        
        #within each MRN group selects the observation with the fewest numbers of NA
        filter(n == min(n)) %>%
        
        #for each MRN, select the first entry
        slice(1) %>%
        
        ungroup() %>%
        
        #drops the added column with number of na's
        select(-n)
```

Some patients when performing match in Duke DEDUCE database were flagged as having
gotten a new MRN. mrn.df contains a list of the old MRNs matched to the new ones.
However, 135 of the old MRNs exist in the supposedly new MRNs which we think are
due to that they had an ecg and echo exam both for the old MRN and the new MRN,
resulting in duplicates after the old MRN gets updated to the new MRN in our dataset.

Our solution to this is:

- identify which MRNs exist both as old and new, exclude these
- update the remaining old MRNs with the new numbers for further patient data mathing
which uses the updated MRNs.

```{r}
#Reads in updated MRN data
mrn.df <- read.csv(file.path(data_dir, "changedMrns.csv"), 
                   stringsAsFactors = FALSE,
                   strip.white = TRUE)

#sets columnnames
colnames(mrn.df) <- c("MRN", "new.MRN")

```

There were duplicates of the new MRN were different old MRNs had been assigned
the same new MRN, all of these are removed below.

```{r}
#selects all rows in mrn.df were new.MRN is not one of the new.MRNs that
#are duplicated.
mrn.df <- mrn.df[!mrn.df$new.MRN %in% 
                         mrn.df[duplicated(mrn.df$new.MRN),]$new.MRN,]

#Excludes all MRNs in df that also exist in mrn.df$new.MRN
df <- df[!df$MRN %in% mrn.df$new.MRN,]

#Return all rows from x, and all columns from x and y.
#Rows in x with no match in y will have NA values in the new columns.
#If there are multiple matches between x and y, all combinations of 
#the matches are returned.
df <- left_join(df, mrn.df, by = "MRN")

#Replaces old MRNs with new ones, dropping the newly added new.MRN column.
df <- df %>%
        mutate(MRN = if_else(!is.na(new.MRN), new.MRN, MRN)) %>%
        select(- new.MRN)
```

## Demographics 

Reading in demographics datas regarding DOB, gender, race and date of death

```{r}
demo.df <- read.csv(file.path(data_dir, "DOB_GENDER_Race_DOD.csv"),
                    stringsAsFactors = FALSE,
                    na.strings=c(""," ","  ", "NA", "N/A"),
                    strip.white = TRUE)

```

Renaming column variables, setting date columns in POSIXct format, (POSIXlt uses a list format, while ct uses time in seconds from the origin, which is more memory effective), converting ordinal variables to factors.

```{r}
demo.df <- demo.df %>%
        transmute(MRN = Duke.MRN, 
               Gender = as.factor(Patient.Gender), 
               Race = as.factor(Patient.Race), 
               DOB = as.POSIXct(Patient.Date.of.Birth, 
                                format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
               Age.at.ECG = as.integer(0), #for future use
               Death.index = as.factor(Patient.Death.Indicator),
               DOD = as.POSIXct(Patient.Death.Date, 
                                format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```


Let's clean up the Race factor levels a bit.

```{r}
levels(demo.df$Race) <- list("Mix"              = c("2 OR MORE RACES", 
                                                    "MULTIRACIAL"),
                             "American.native"  = c("ALASKAN NATIVE",
                                                     "AMERICAN INDIAN",
                                                     "AMERICAN INDIAN OR ALASKAN NATIVE"),
                             "Native.hawaiian.or.other.pacific.islander" = 
                               "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER",
                             "Caucasian"        = c("WHITE OR CAUCASIAN", 
                                                    "CAUCASIAN/WHITE"),
                             "African.american" = "BLACK OR AFRICAN AMERICAN",
                             "NA"               = c("UNAVAILABLE", 
                                                    "NOT REPORTED/DECLINED"),
                             "Asian"            = "ASIAN",
                             "Other"            = "OTHER")
```

## Matching demographics

This way of merging returns all rows from df where there are matching values in demo.df, and all columns from df and demo.df. If there are multiple matches between df and demo.df, all combination of the matches are returned.

```{r}
main.df <- inner_join(df, demo.df, by = "MRN") %>%
        
        #we need the patient age at the time of the ecg
        mutate(Age.at.ECG = round(as.integer(difftime(ECGDate, 
                                                      DOB, 
                                                      units = "days"))/365.25))
        
```

Read in the Cardiac Resynchronization Therapy (CRT) data.

```{r}
crt.df <- read.csv(file.path(data_dir, "Dx/CRT Codes.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

crt.df <- crt.df %>%
        rename(MRN = Duke.MRN, 
               CRT.Date = ICD.CPT.Procedure.Date) %>%
        mutate(CRT.Date = as.POSIXct(CRT.Date, 
                                    format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
               CRT = as.factor(1),
               Censor = NA) %>%
        group_by(MRN) %>%
        filter(CRT.Date == min(CRT.Date)) %>%
        slice(1) %>%
        ungroup()
```

Read in LVAD data

```{r}
lvad.df <- read.csv(file.path(data_dir, "Dx/LVAD Codes.csv"),
                    stringsAsFactors = FALSE,
                    na.strings=c(""," ","  ", "NA", "N/A"))

lvad.df <- lvad.df %>%
        rename(MRN = Duke.MRN,
               LVAD.Code = ICD.Procedure.Code,
               LVAD.Date = Procedure.Date) %>%
        mutate(LVAD.Date = as.POSIXct(LVAD.Date, 
                                                format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
               LVAD = as.factor(1)) %>%
        group_by(MRN) %>%
        filter(LVAD.Date == min(LVAD.Date)) %>%
        slice(1) %>%
        ungroup()
```


Read in Heart transplant data

```{r}
htx.df <- read.csv(file.path(data_dir, "Dx/Hearttransplant Codes.csv"),
                    stringsAsFactors = FALSE,
                    na.strings=c(""," ","  ", "NA", "N/A"))

htx.df <- htx.df %>%
        rename(MRN = Duke.MRN, 
               HTx.Code = ICD.Procedure.Code,
               HTx.Date = Procedure.Date) %>%
        mutate(HTx.Date = as.POSIXct(HTx.Date, 
                                                format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
               HTx = as.factor(1)) %>%
  group_by(MRN) %>%
  filter(HTx.Date == min(HTx.Date)) %>%
  slice(1) %>%
  ungroup()

```


Read in heart failure data.

```{r}
#Read heart failure data
hf.df <- read.csv(file.path(data_dir, "Dx/HF DX.csv"), 
                  stringsAsFactors = FALSE,
                  na.strings=c(""," ","  ", "NA", "N/A"))

#Renames HF variables
hf.df <- hf.df %>%
        rename(MRN = Duke.MRN, 
               HF.Code = ICD.Diagnosis.Code,
               HF.Date = Diagnosis.Date) %>%
        mutate(HF.Date = as.POSIXct(HF.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))

```

Read in cad data.

```{r}

#Read CAD data
cad.df <- read.csv(file.path(data_dir, "Dx/CAD DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

#Renames CAD variables
cad.df <- cad.df %>%
        rename(MRN = Duke.MRN, 
               CAD.Code = ICD.Diagnosis.Code, 
               CAD.Date = Diagnosis.Date) %>%
        mutate(CAD.Date = as.POSIXct(CAD.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Read in OSA data.

```{r}
osa.df <- read.csv(file.path(data_dir, "Dx/OSA DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

osa.df <- osa.df %>%
        rename(MRN = Duke.MRN, 
               OSA.Code = ICD.Diagnosis.Code, 
               OSA.Date = Diagnosis.Date) %>%
        mutate(OSA.Date = as.POSIXct(OSA.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Read in CHD data.

```{r}
chd.df <- read.csv(file.path(data_dir, "Dx/CHD DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

chd.df <- chd.df %>%
        rename(MRN = Duke.MRN, 
               CHD.Code = ICD.Diagnosis.Code, 
               CHD.Date = Diagnosis.Date) %>%
        mutate(CHD.Date = as.POSIXct(CHD.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Read in COPD data.

```{r}
copd.df <- read.csv(file.path(data_dir, "Dx/COPD DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

copd.df <- copd.df %>%
        rename(MRN = Duke.MRN, 
               COPD.Code = ICD.Diagnosis.Code, 
               COPD.Date = Diagnosis.Date) %>%
        mutate(COPD.Date = as.POSIXct(COPD.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Read in DM data.

```{r}
dm.df <- read.csv(file.path(data_dir, "Dx/DM DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

dm.df <- dm.df %>%
        rename(MRN = Duke.MRN, 
               DM.Code = ICD.Diagnosis.Code, 
               DM.Date = Diagnosis.Date) %>%
        mutate(DM.Date = as.POSIXct(DM.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Read in HTN data.

```{r}
htn.df <- read.csv(file.path(data_dir, "Dx/HTN DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

htn.df <- htn.df %>%
        rename(MRN = Duke.MRN, 
               HTN.Code = ICD.Diagnosis.Code, 
               HTN.Date = Diagnosis.Date) %>%
        mutate(HTN.Date = as.POSIXct(HTN.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```

Reading in AF data

```{r}
af.df <- read.csv(file.path(data_dir, "Dx/AF DX.csv"), 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A"))

af.df <- af.df %>%
        rename(MRN = Duke.MRN, 
               AF.Code = ICD.Diagnosis.Code, 
               AF.Date = Diagnosis.Date) %>%
        mutate(AF.Date = as.POSIXct(AF.Date, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"))
```


Creating new dataframe containing just the MRNs and the corresponding ECGdates
so that this can be used in the dx.match function to match diagnosis date to ecg date.

```{r}
ecgdates.df <- main.df %>%
        select(MRN, ECGDate)
```

Function to select diagnosis meeting match criteria, the first date that is not more than >30 days after ECG date.

```{r}
dx.match.fun <- function(x, diagnosis) { 
  
        #x should be dataframe, diagnosis in quotes
        #and capital letters
  
        #Creates date variable with the diagnosis in capital letters
        #followed by date
        date <- paste(diagnosis, "Date", sep = ".")
        
        #filters for the earliest diagnosis date
        filter_call <- interp(~ a == min(a) , a = as_name(date))
        
        #filters for the diagnosis date that are not more than 30 days after
        #ecgdate
        filter_call2 <- interp(~ as.numeric(difftime(ECGDate, a, units = "days")) >=(-30),
                               a = as_name(date))
        
        #starts with joining x with ecgdates dataframe
        inner_join(x, ecgdates.df, by = "MRN") %>%
                group_by(MRN) %>%
                filter_(.dots = list(filter_call)) %>%
                filter_(.dots = list(filter_call2)) %>%
                
                #for some patients there will be multiple entries with the
                #same diagnosis date, this is due to different
                #diagnosis codes, which one does not matter
                #so we are selecting the first one
                slice(1) %>%
                
                ungroup() %>%
                
                #creates a new column with logical indicating presence of the current diagnosis
                mutate_(.dots = setNames(list(~ as.factor(1)), diagnosis)) %>%
                
                select(-ECGDate)
}
```

```{r}
af.df <- dx.match.fun(af.df, "AF")

cad.df <- dx.match.fun(cad.df, "CAD")

chd.df <- dx.match.fun(chd.df, "CHD")

copd.df <- dx.match.fun(copd.df, "COPD")

dm.df <- dx.match.fun(dm.df, "DM")

hf.df <- dx.match.fun(hf.df, "HF")

htn.df <- dx.match.fun(htn.df, "HTN")

osa.df <- dx.match.fun(osa.df, "OSA")

```

Joining diagnosis data to main.df

```{r}

main.df2 <- left_join(main.df, af.df, by = "MRN") %>%
        left_join(cad.df, by = "MRN") %>%
        left_join(chd.df, by = "MRN") %>%
        left_join(copd.df, by = "MRN") %>%
        left_join(crt.df, by = "MRN") %>%
        left_join(dm.df, by = "MRN") %>%
        left_join(hf.df, by = "MRN") %>%
        left_join(htn.df, by = "MRN") %>%
        left_join(htx.df, by = "MRN") %>%
        left_join(lvad.df, by = "MRN") %>%
        left_join(osa.df, by = "MRN")

```

```{r}
# Exclude all patients with an endpointdate less than 24 hours after ecgdate.

endPointAfterEcg <- function(ecgdate, endpointdate) {
  x <- (as.numeric(difftime(ecgdate, endpointdate, units = "days")) <= (-1))
  
  return(x)
}


main.df2 <- main.df2 %>%
  group_by(MRN) %>%
  filter((endPointAfterEcg(ECGDate, CRT.Date) | is.na(CRT.Date))) %>%
  filter((endPointAfterEcg(ECGDate, LVAD.Date) | is.na(LVAD.Date))) %>%
  filter((endPointAfterEcg(ECGDate, HTx.Date) | is.na(HTx.Date))) %>%
  ungroup()
```


## Medications

```{r}
read.rx.data <- function(filepath, rx) {
        
        rx.name <- paste(rx, "Rx.Name", sep = ".")
        rx.start <- paste(rx, "Rx.Start", sep = ".")
        rx.end <- paste(rx, "Rx.End", sep = ".")
        
        mutate_rx.start <- interp(~ as.POSIXct(a, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
                              a = as_name(rx.start))
        
        mutate_rx.end <- interp(~ as.POSIXct(a, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
                               a = as_name(rx.end))
        
        read.csv(file = filepath, 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A")) %>%
                
                rename_(.dots = setNames(list("Duke.MRN",
                                      "Medication.Name",
                                      "Start.Date",
                                      "End.Date"),
                                 list("MRN",
                                      rx.name,
                                      rx.start,
                                      rx.end))) %>%
                mutate_(.dots = setNames(list(mutate_rx.start,
                                              mutate_rx.end), 
                                         list(rx.start,
                                              rx.end)))
}
```

```{r}
classIII.df <- read.rx.data(file.path(data_dir, "Medications/Class III Patient Medications.csv"),
                            "ClassIII")
```

```{r}
classIc.df <- read.rx.data(file.path(data_dir, "Medications/Class Ic Patient Medication data.csv"),
                           "ClassIc")
```

```{r}
bb1.df <- read.rx.data(file.path(data_dir, "Medications/BB Patient Med2013 to 2017.csv"),
                           "BB")
```

```{r}
bb2.df <- read.rx.data(file.path(data_dir, "Medications/BB Patient Med2010 to 2013.csv"),
                           "BB")
```

```{r}
bb3.df <- read.rx.data(file.path(data_dir, "Medications/BB Patient Med2006 to 2010.csv"),
                           "BB")
```

```{r}
bb.df <- rbind(bb1.df, bb2.df, bb3.df)
```

```{r}
amiodarone.df <- read.rx.data(file.path(data_dir, "Medications/AmiodaronePatient Medication data.csv"),
                           "Amiodarone")
```

```{r}
acei1.df <- read.rx.data(file.path(data_dir, "Medications/ACEI Patient Medications 2013to2017.csv"),
                           "ACEI")
```

```{r}
acei2.df <- read.rx.data(file.path(data_dir, "Medications/ACEI Patient Medications 2010to2013.csv"),
                           "ACEI")
```

```{r}
acei3.df <- read.rx.data(file.path(data_dir, "Medications/ACEI Patient Medications 2006to2010.csv"),
                           "ACEI")
```

```{r}
acei.df <- rbind(acei1.df, acei2.df, acei3.df)
```

```{r}
rx.match.fun <- function(x, rx) {
      
        startdate <- paste(rx, "Rx.Start", sep = ".")
        
        enddate <- paste(rx, "Rx.End", sep = ".")
        
        rx.log <- paste(rx, "Rx", sep = ".")
        
        #filters for the Rx startdate that is not more than 30 days after
        #ecgdate and Rx enddate not more than 30 days before ecgdate
        date_filter <- interp(~ as.numeric(difftime(ECGDate, a, units = "days")) >=(-30) &
                                           as.numeric(difftime(ECGDate, b, units = "days")) <=(30),
                                   a = as_name(startdate),
                                   b = as_name(enddate))
        
        #starts with joining x with ecgdates dataframe
        inner_join(x, ecgdates.df, by = "MRN") %>%
                group_by(MRN) %>%
                filter_(.dots = list(date_filter)) %>%
                
                #adds column indicating presence of type of medication
                mutate_(.dots = setNames(list(~as.factor(1)), rx.log)) %>%
                slice(1) %>%
                ungroup() %>%
                select(-ECGDate)
                
}
```

```{r}
classIII.df <- rx.match.fun(classIII.df, "ClassIII")

classIc.df <- rx.match.fun(classIc.df, "ClassIc")

acei.df <- rx.match.fun(acei.df, "ACEI")

amiodarone.df <- rx.match.fun(amiodarone.df, "Amiodarone")

bb.df <- rx.match.fun(bb.df, "BB")
```

```{r}
read.rx.omr.data <- function(filepath, rx) {
        
        rx.name <- paste(rx, "Rx.OMR.Name", sep = ".")
        rx.date <- paste(rx, "Rx.OMR.Date", sep = ".")
        
        mutate_rx.date <- interp(~ as.POSIXct(a, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
                              a = as_name(rx.date))
        
        read.csv(file = filepath, 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A")) %>%
                
                rename_(.dots = setNames(list("Duke.MRN",
                                      "OMR.Medication.Name",
                                      "OMR.Report.Date"),
                                 list("MRN",
                                      rx.name,
                                      rx.date))) %>%
                
                mutate_(.dots = setNames(list(mutate_rx.date),rx.date))
}
```

```{r}
classIII.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/Class III OMR data.csv"),
                           "ClassIII")
```

```{r}
classIc.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/Class Ic OMR data.csv"),
                           "ClassIc")
```

```{r}
bb1.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/BB OMR2010 to 2013.csv"),
                           "BB")
```

```{r}
bb2.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/BB OMR2006 to 2010.csv"),
                           "BB")
```

```{r}
bb.OMR.df <- rbind(bb1.OMR.df, bb2.OMR.df)
```

```{r}
amiodarone.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/AmiodaroneOMR data.csv"),
                           "Amiodarone")
```

```{r}
acei1.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/ACEI OMR2010 to 2013.csv"),
                           "ACEI")
```

```{r}
acei2.OMR.df <- read.rx.omr.data(file.path(data_dir, "Medications/ACEI OMR2006 to 2010.csv"),
                           "ACEI")
```

```{r}
acei.OMR.df <- rbind(acei1.OMR.df, acei2.OMR.df)
```


```{r}
rx.omr.match <- function(x, rx) {
        rxdate <- paste(rx, "Rx.OMR.Date", sep = ".")
        
        mutate_difftime <- interp(~ abs(as.numeric(difftime(ECGDate, a, units = "days"))),
                                  a = as_name(rxdate))
        
        #filters for the Rx date that is within 30 days of
        #ecgdate
        rxdate_filter <- interp(~ abs(as.numeric(difftime(ECGDate, a, units = "days"))) <=30,
                                   a = as_name(rxdate))
        
        #starts with joining x with ecgdates dataframe
        inner_join(x, ecgdates.df, by = "MRN") %>%
                group_by(MRN) %>%
                mutate_(.dots = setNames(list(mutate_difftime), "diff")) %>%
                filter(diff == min(diff)) %>%
                filter_(.dots = rxdate_filter) %>%
                slice(1) %>%
                ungroup() %>%
                select(-c(diff,ECGDate))
}
```

```{r}
classIII.OMR.df <- rx.omr.match(classIII.OMR.df, "ClassIII")
```

```{r}
classIc.OMR.df <- rx.omr.match(classIc.OMR.df, "ClassIc")
```

```{r}
bb.OMR.df <- rx.omr.match(bb.OMR.df, "BB")
```

```{r}
amiodarone.OMR.df <- rx.omr.match(amiodarone.OMR.df, "Amiodarone")
```

```{r}
acei.OMR.df <- rx.omr.match(acei.OMR.df, "ACEI")
```

Joining diagnosis data to main.df
Earlier added column indicating whether specific type of medication was used or not
except from joining hopstial and OMR patient medcation data this code also updates
that column with a 1 if presence of entry in OMR medication column

```{r}
main.df3 <- left_join(main.df2, classIII.df, by = "MRN") %>%
        
        left_join(classIII.OMR.df, by = "MRN") %>%
        
        mutate(ClassIII.Rx = ifelse(!is.na(ClassIII.Rx.OMR.Name), 1, ClassIII.Rx)) %>%

        left_join(classIc.df, by = "MRN") %>%
        
        left_join(classIc.OMR.df, by = "MRN") %>%
        
        mutate(ClassIc.Rx = ifelse(!is.na(ClassIc.Rx.OMR.Name), 1, ClassIc.Rx)) %>%
        
        left_join(acei.df, by = "MRN") %>%
        
        left_join(acei.OMR.df, by = "MRN") %>%
        
        mutate(ACEI.Rx = ifelse(!is.na(ACEI.Rx.OMR.Name), 1, ACEI.Rx)) %>%
        
        left_join(amiodarone.df, by = "MRN") %>%
        
        left_join(amiodarone.OMR.df, by = "MRN") %>%
        
        mutate(Amiodarone.Rx = ifelse(!is.na(Amiodarone.Rx.OMR.Name), 1, Amiodarone.Rx)) %>%
        
        left_join(bb.df, by = "MRN") %>%
        
        left_join(bb.OMR.df, by = "MRN") %>%
        
        mutate(BB.Rx = ifelse(!is.na(BB.Rx.OMR.Name), 1, BB.Rx))

```

## Lab data

```{r}
read.lab.data <- function(filepath, lab) {
        
        lab.date <- paste(lab, "lab.Date", sep = ".")
        lab.value <- paste(lab, "lab.Value", sep = ".")
        
        mutate_lab.date <- interp(~ as.POSIXct(a, format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York"),
                              a = as_name(lab.date))
        
        #if there are Na's in the text result value these are replced with the value
        #in the numeric result column
        mutate_lab.value <- interp(~ifelse(is.na(a), Numeric.Result,a),
                                   a = as_name(lab.value))
        
        #all entries with NA in the value column are excluded 
        filter_lab.value <- interp(~ !is.na(a),
                                   a = as_name(lab.value))
        
        read.csv(file = filepath, 
                   stringsAsFactors = FALSE,
                   na.strings=c(""," ","  ", "NA", "N/A")) %>%
                
                rename_(.dots = setNames(list("Duke.MRN",
                                      "Specimen.Collection.Date",
                                      "Text.Result"),
                                 list("MRN",
                                      lab.date,
                                      lab.value))) %>%
                
                mutate_(.dots = setNames(list(mutate_lab.date),lab.date)) %>%
                mutate_(.dots = setNames(list(mutate_lab.value),
                                         lab.value)) %>%
                select(-c(Test.Description, Numeric.Result)) %>%
                filter_(.dots = list(filter_lab.value))
}
```

```{r}
na1.df <- read.lab.data(file.path(data_dir, "Lab/NA2013_2016.csv"), "Na")
na2.df <- read.lab.data(file.path(data_dir, "Lab/NA2010_2013.csv"), "Na")
na3.df <- read.lab.data(file.path(data_dir, "Lab/NA2006_2010.csv"), "Na")
na4.df <- read.lab.data(file.path(data_dir, "Lab/NA2001_2006.csv"), "Na")

na.df <- rbind(na1.df, na2.df, na3.df, na4.df)
```

```{r}
k1.df <- read.lab.data(file.path(data_dir, "Lab/K2013_2016.csv"), "K")
k2.df <- read.lab.data(file.path(data_dir, "Lab/K2010_2013.csv"), "K")
k3.df <- read.lab.data(file.path(data_dir, "Lab/K2006_2010.csv"), "K")
k4.df <- read.lab.data(file.path(data_dir, "Lab/K2001_2006.csv"), "K")

k.df <- rbind(k1.df, k2.df, k3.df, k4.df)
```

```{r}
gfr.df <- read.csv(file.path(data_dir, "Lab/GFR DX.csv"),
                   stringsAsFactors = FALSE,
                   strip.white = TRUE,
                   na.strings = c(""," ","  ", "NA", "N/A")) %>%
        
        rename(MRN = Duke.MRN) %>%
       
        mutate(Specimen.Collection.Date = as.POSIXct(Specimen.Collection.Date,
                                                     format = "%m/%d/%Y %H:%M")) %>%
       
        mutate(Text.Result = ifelse(is.na(Text.Result),Numeric.Result,Text.Result)) %>%
        
        select(-c(Test.Description, Numeric.Result)) %>%
        
        filter(!is.na(Text.Result)) %>%
        
        rename(GFR.lab.Date = Specimen.Collection.Date,
               
               GFR.lab.Value = Text.Result)
```


```{r}
bnp1.df <- read.csv(file.path(data_dir, "Lab/BNP AND PRO BNP2.csv"),
                   stringsAsFactors = FALSE,
                   strip.white = TRUE,
                   na.strings = c(""," ","  ", "NA", "N/A")) %>%
        rename(MRN = Duke.MRN) %>%
        mutate(Specimen.Collection.Date = as.POSIXct(Specimen.Collection.Date,
                                                     format = "%m/%d/%Y %H:%M:%S", tz = "America/New_York")) %>%
        mutate(Text.Result = ifelse(is.na(Text.Result),Numeric.Result,Text.Result)) %>%
        select(-Numeric.Result) %>%
        filter(!is.na(Text.Result))

bnp1.df$Test.Description <- as.factor(bnp1.df$Test.Description)
levels(bnp1.df$Test.Description) <- list(
        "BNP" = c("BNP - LABCORP",
                  "BRAIN NATRIURETIC PEPTIDE",
                  "BRAIN NATRIURETIC PEPTIDE     (BKR)",
                  "BTYPE NATRIURETIC PEPTIDE DUAP"),
        "ProBNP" = c("PRO BRAIN NATRIURETIC PEPTIDE",
                     "PRO BRAIN NATRIURETIC PEPTIDE     (BKR)",
                     "PROBNP - LABCORP"))
```

```{r}
bnp.df <- bnp1.df %>%
        filter(Test.Description == "BNP") %>%
        rename(BNP.lab.Date = Specimen.Collection.Date,
               BNP.lab.Value = Text.Result) %>%
        select(-Test.Description)

probnp.df <- bnp1.df %>%
        filter(Test.Description == "ProBNP") %>%
        rename(ProBNP.lab.Date = Specimen.Collection.Date,
               ProBNP.lab.Value = Text.Result) %>%
        select(-Test.Description)
```


```{r}
lab.match.fun <- function(x, lab) {
        labdate <- paste(lab, "lab.Date", sep = ".")
        labvalue <- paste(lab, "lab.Value", sep = ".")
        labvalue.x <- paste(lab, "lab.Value.x", sep = ".")
        
        #measures the time difference in days between ecgdate and labdate
        mutate_difftime <- interp(~ abs(as.numeric(difftime(ECGDate, a, units = "days"))),
                                  a = as_name(labdate))
        
        #filters for the lab date that is within 30 days of
        #ecgdate
        labdate_filter <- interp(~ abs(as.numeric(difftime(ECGDate, a, units = "days"))) <=30,
                                   a = as_name(labdate))
        
        #removes characters that prevent coercion to numerical values
        mutate_labvalue <- interp(~ str_replace_all(a, "[,= \\*<>\\(GCANCELED\\%]",""),
                                               a = as_name(labvalue))
        
        #coerces to numerical values
        mutate_labvalue2 <- interp(~ as.numeric(a), a = as_name(labvalue.x))
        
        #excludes all entries with NA in the labvalue column
        filter_labvalue.x <- interp(~ !is.na(a), a = as_name(labvalue.x))
        
        #starts with joining x with ecgdates dataframe
        inner_join(x, ecgdates.df, by = "MRN") %>%
                
                group_by(MRN) %>%
                
                mutate_(.dots = setNames(list(mutate_difftime), "diff")) %>%
                
                #filters for the labvalue closest in time to ecgdate
                filter(diff == min(diff)) %>%
                
                filter_(.dots = labdate_filter) %>%
                
                slice(1) %>%
                
                ungroup() %>%
                
                mutate_(.dots = setNames(list(mutate_labvalue), labvalue.x)) %>%
                
                mutate_(.dots = setNames(list(mutate_labvalue2), labvalue.x)) %>%
                
                select(-c(diff,ECGDate)) %>%
                
                filter_(.dots = filter_labvalue.x)
}
```

```{r}
na.df <- lab.match.fun(na.df, "Na")
k.df <- lab.match.fun(k.df, "K")
bnp.df <- lab.match.fun(bnp.df, "BNP")
probnp.df <- lab.match.fun(probnp.df, "ProBNP")
gfr.df <- lab.match.fun(gfr.df, "GFR")
```

```{r}
main.df4 <- left_join(main.df3, na.df, by = "MRN") %>%
        left_join(k.df, by = "MRN") %>%
        left_join(bnp.df, by = "MRN") %>%
        left_join(probnp.df, by = "MRN") %>%
        left_join(gfr.df, by = "MRN")
```

## Data processing

```{r}
#drops unused factor levels
main.df5 <- droplevels(main.df4)
```

### Exclusion criteria

- Patients under 18 years of age
- QRSduration <60ms

```{r}

main.df6 <- main.df5 %>%
        
        filter(Age.at.ECG >=18,
               
               QRSDur >=60)
```

```{r}
main.df6 <- main.df6 %>%
        mutate(RBBB = if_else(RBBB == 1 & QRSDur >= 120,
                               true  = 1,
                               false = 0),
               
               RBBB_LAFB = if_else(RBBB == 1 & LAFB == 1,
                                    true  = 1,
                                    false = 0),
               
               RBBB_LPFB = if_else(RBBB == 1 & LPFB == 1,
                                    true  = 1,
                                    false = 0),
               
               RBBB = if_else(RBBB_LAFB == 1 | RBBB_LPFB == 1,
                               true  = 0,
                               false = RBBB),
               
               LBBB = if_else(LBBB == 1 & QRSDur >= 120,
                               true  = 1,
                               false = 0),
               
               IVCD = if_else(IVCD == 1 & QRSDur >= 110 &
                                       LBBB == 0 &
                                       RBBB == 0 &
                                       RBBB_LAFB == 0 &
                                       RBBB_LPFB == 0,
                               true  = 1,
                               false = 0),
               
               LAFB = if_else(LAFB == 1 & QRSDur <120 & IVCD == 0,
                               true  = 1,
                               false = 0),
               
               LPFB = if_else(LPFB == 1 & QRSDur <120 & IVCD == 0,
                               true  = 1,
                               false = 0),
               
               Normal_conduction = if_else(LBBB      == 0 & 
                                           LAFB      == 0 &
                                           LPFB      == 0 &
                                           RBBB      == 0 &
                                           RBBB_LAFB == 0 &
                                           RBBB_LPFB == 0 &
                                           IVCD      == 0 &
                                           PACE      == 0 &
                                           QRSDur < 120,
                                           
                                           true  = 1,
                                           false= 0))
               
               
```

```{r}
main.df6 <- droplevels(main.df6)
```

```{r}
replace_na <- function(x) {
        x[is.na(x)] <- 0
        x
}
```

```{r}
fix_na_columns <- c("AF", 
                    "CAD", 
                    "CHD",
                    "COPD",
                    "CRT",
                    "DM",
                    "HF",
                    "HTN",
                    "HTx",
                    "LVAD",
                    "OSA",
                    "ClassIII.Rx",
                    "ClassIc.Rx",
                    "ACEI.Rx",
                    "Amiodarone.Rx",
                    "BB.Rx")
```


```{r}
main.df6 <- main.df6 %>%
        mutate_at(fix_na_columns,as.numeric) %>%
        mutate_at(fix_na_columns,replace_na)
```

```{r}
# patients without ecg classification or multiple entries are excluded
promise <- main.df6 %>%
        rowwise() %>%
        mutate(n_entries = sum(Normal_conduction,
                               IVCD,
                               LBBB,
                               LAFB,
                               LPFB,
                               RBBB,
                               RBBB_LAFB,
                               RBBB_LPFB,
                               PACE)) %>%
        filter(n_entries == 1) %>%
        select(-n_entries) %>%
        ungroup()
```